Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over \cye{convex domains and} domains with shellable triangulations

### 2024-09-27 revision/main.tex
include MV revision in a subfolder

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
% \usepackage{fullpage}
\input{header}

@@

% Title, authors, and affiliations
% \title{Poincar\'e--Friedrichs constants over domains with shellable triangulations}
% \author{Th\'eophile Chaumont-Frelet, Martin Werner Licht, Martin Vohral\'ik}

\title{Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations}
\author{
    Th\'eophile Chaumont-Frelet\thanks{Affiliation of the First Author, \texttt{email1@example.com}} \and
    Martin Werner Licht\thanks{Affiliation of the Second Author, \texttt{martin.licht@epfl.ch}} \and
    Martin Vohral\'ik\thanks{Affiliation of the Third Author, \texttt{email3@example.com}}
}

@@

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\notice}[1]{\textcolor{red}{REMARK: #1}}

@@

\begin{abstract}
    We construct potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over domains with shellable triangulations. This class of triangulations includes local stars in dimension two and three. 
    The operator norms of our potentials satisfy explicitly computable bounds that only depend on the geometry. 
    We thus obtain computable upper bounds for Poincar\'e--Friedrichs inequalities and computable lower bounds for the eigenvalues of several vector Laplacians. 
    % 
    %     We construct upper bounds for Poincar\'e--Friedrichs constants over triangulated domains. 
    %     This includes lower estimates for the singular values of the gradient, curl, and divergence operators as special cases. 
    %     The main application that we envision are computable estimates over finite element patches. 
    %
    As a result with independent standing, we establish Poincar\'e--Friedrichs inequalities for the $L^{p}$ de~Rham complex over convex bounded domains. 
    This is achieved through bounding operator norms of Poincar\'e and Bogovski\u{\i} potential operators.
    We express all our main results in the calculus of differential forms and treat the gradient, curl, and divergence operator as its particular instances. 
\end{abstract}

@@

% DONE {Reformulate the first paragraph. } % 
Potentials for the differential operators of vector calculus and exterior calculus are of fundamental importance.
The operator norms of the potentials are upper bounds for the Poincar\'e--Friedrichs constants, 
which quantify the fundamental stability features of a variety of partial differential equations.
% which serve a variety of purposes:
% these constants quantify the fundamental stability properties of partial differential equations, imply lower bounds for the spectra of vector and Hodge Laplacians, and they . 
However, while potential operators and Poincar\'e--Friedrichs constants for the gradient have been subject to extensive study, quantifiable results regarding the curl and divergence operators or the exterior derivative are not nearly as numerous yet. 
% We study potential operators in the special class of domains with shellable triangulations,
% which include local stars in 2D and 3D. 
% We bound the operator norms of our operators in terms of the shape regularity of the triangulation.

% We contribute to the construction of potential operators in vector analysis and exterior calculus over triangulated domains. 
% We address the construction of potentials for the gradient, the curl and the divergence operators, and more generally the exterior derivative, over triangulated domains. 
% The operator norms of these potentials will satisfy explicitly computable bounds in Lebesgue norms only in terms of the geometry. 
% In particular, such operator norms bound yield computable upper bounds for Poincar\'e--Friedrichs inequalities and computable lower bounds for the eigenvalues of several vector Laplacians.

% DONE : MV : remove 'For example' 
We illustrate the subject of our discussion with the curl operator. 
% One example application concerns the curl operator. 
If $\bff \in \bfL^{2}(\Omega)$ is a square-integrable vector field on some domain $\Omega \subseteq \bbR^3$, 
we want to find a potential $\bfu \in \bfL^2(\Omega)$ under the curl operator, i.e., $\curl \bfu = \bff$, such that the $\bfL^2$ norm of the potential is bounded in terms of the source field $\bff$. Specifically, we want the Poincar\'e--Friedrichs inequality % DONE: Theo prefers not using ``expressed''
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfL^{2}(\Omega) \\ \curl \bfv = \bff } } \| \bfv \|_{L^{2}(\Omega)}
    \leq 
    C_{\curl}
    \| \bff \|_{L^{2}(\Omega)}
    .
\end{align*}
The constant in this inequality corresponds to a lower bound for the Maxwell eigenvalues and quantifies the stability properties of the Maxwell system on that domain. 
%\color{blue}
Broadly speaking, upper bounds for the Poincar\'e--Friedrichs constants correspond to lower bounds for the non-zero eigenvalues of the scalar and vector Laplacians.
% Furthermore, their constants enter the stability and convergence theory of finite element methods. 
In numerical methods, estimates for the Poincar\'e--Friedrichs constant enter the stability and convergence theory of finite element methods, as well as spectral bounds for finite element matrices. 
% \color{black}
\\

Finding a potential operator, whether linear or non-linear, and bounding the Poincar\'e--Friedrichs constant is generally not an easy undertaking. 
The majority of estimates in the literature address only gradient potentials. 
% We accomplish our results for a special class of triangulations, so-called shellable triangulations.
% We accomplish this task in 
We carry out this endeavor in the special case where $\Omega$ admits a so-called \emph{shellable} triangulation. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards the more general case. 
For example, local stars around a simplex within a larger 2D or 3D triangulation are shellable.
Poincar\'e--Friedrichs inequalities over local stars are our main goal in this exposition. 
\\

Before we outline the main results of this exposition in more detail, 
we first review the state of the literature on Poincar\'e--Friedrichs inequalities on specific classes of domains. 
We take the Poincar\'e--Friedrichs inequality for the gradient of scalar functions as our conceptual point of reference.

% DONE: MV : Be more didactical: write following as the existence of a potential. 
Given a domain $\Omega \subseteq \bbR^{n}$ and Lebesgue exponent $1 \leq p \leq \infty$,
we search for a constant $C_{\grad,\Omega,p} > 0$ such that the following holds:
for every gradient vector field $\bff \in \nabla W^{1,p}(\Omega)$ there exists a scalar potential $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:intro:inequalitypf:gradient} 
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
We call this inequality \emph{Poincar\'e--Friedrichs inequality} and the constant $C_{\grad,\Omega,p}$ is called the \emph{Poincar\'e--Friedrichs constant} with exponent $p$. 
The different scalar potentials only differ by a constant. 
Therefore, %when the domain is connected, then 
this inequality is equivalent to the following statement: 
\begin{align}\label{math:intro:poincarefriedrichs:grad}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
    \quad 
    \forall
    u \in W^{1,p}(\Omega) % DONE: Use \forall notation 
    .\tag{PF} 
\end{align}
% Given a domain $\Omega \subseteq \bbR^{n}$ and Lebesgue exponent $1 \leq p \leq \infty$,
% we search for a constant $C_{\grad,\Omega,p} > 0$ such that 
% \begin{align}\label{math:intro:poincarefriedrichs:grad}
%     \min\limits_{c\in\bbR}
%     \| u - c \|_{L^{p}(\Omega)}
%     \leq 
%     C_{\grad,\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
%     \quad 
%     u \in W^{1,p}(\Omega) % DONE: Use \forall notation 
%     .\tag{PF} 
% \end{align}
Finding a norm-minimizing potential of a given gradient $\nabla u$, if possible, is generally a nonlinear operation.\footnote{The existence of a minimum-norm potential is not guaranteed unless $1 < p < \infty$.}
The operator norm of any \emph{linear} potential construction serves as an upper bound for the Poincar\'e--Friedrichs constant. 
% Even though the existence of such a bounded linear potential construction is not evident a priori, picking the average-free potential 
One very simple example for such a bounded linear operator is the average-free potential: % DONE Theo : too much formalism. Solution: Acknowledge that it's very formal 
\begin{align*}
    \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the constant $C_{{\rm P},\Omega,p} > 0$ in the Poincar\'e inequality 
\begin{align}\label{math:intro:poincare:grad}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm P},\Omega,p} \| \nabla u \|_{L^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    , \tag{P} % DONE: tag Poincar\'e inequality 
\end{align}
where $u_\Omega$ denotes the average of $u$.
Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities~\eqref{math:intro:poincarefriedrichs:grad} or Poincar\'e inequalities~\eqref{math:intro:poincare:grad}.

Most research on optimal constants has addressed the case of convex domains. 
% The optimal constants that depend solely on the domain diameter are known if $p=2$ \cite{bebendorf2003note} and if $p=1$ \cite{acosta2004optimal}. 
Notably, if the constants are required to depend on the convex domain only via its diameter, 
then the optimal constants $C_{\grad,\Omega,p}$ for any $1 \leq p < \infty$ are known explicitly~\cite{bebendorf2003note,acosta2004optimal,esposito2013poincare,ferone2012remark}.
When the domain is not convex but star-shaped with respect to a ball,
then polynomial interpolation estimates already imply the Poincar\'e inequality~\cite{brenner2008mathematical,ern2021finite}. 
Poincar\'e inequalities hold over star-shaped open sets~\cite[Theorem~3.1]{hurri1988poincare}. % DONE: strange sentence

The situation is less clear cut when we study the curl or divergence operators in vector calculus.
We use the vector field spaces
\begin{align*}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) : \curl \bfu \in \bfL^{p}(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) : \divergence \bfu \in L^{p}(\Omega) \right\}
    .
\end{align*}
Their members are those vector fields whose distributional curls and divergences are in Lebesgue spaces. 
This only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector fields. 
Our objective is to find bounded potentials for the operators 
\begin{align*}
    \curl : \bfW^{p}(\curl,\Omega) \rightarrow \bfL^{p}(\Omega),
    \qquad 
    \divergence : \bfW^{p}(\divergence,\Omega) \rightarrow L^{p}(\Omega),
\end{align*}
The fundamental difference to the gradient is that the curl and divergence have infinite-dimensional kernels. 
In the Banach space case, it is not immediately evident that the kernels of these operators are complemented.
The existence of a bounded linear potential operator is a non-trivial fact even in the Hilbert space case.
We are interested in constants for inequalities such as 
\begin{align*}
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\curl,\Omega) \\ \curl \bfv = \bff } } 
    \| \bfv \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)}
    ,
    \\%\qquad 
    \min\limits_{ \substack{ \bfv \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfv = f } } 
    \| \bfv \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)}
    ,
\end{align*}
and how these constants can be estimated by the operator norms of linear potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}

Not much attention seems to have been given to the study of optimal constants in such Poincar\'e--Friedrichs inequalities for the curl and divergence operators. 
% even if the domain is convex, 
This sharply contrasts to the study of the gradient operator. 
%
% The literature in functional analysis is much scarcer when it comes to any upper bounds for the Poincar\'e--Friedrichs constants. 
We can access some estimates from the literature in the formalism of exterior calculus. 
We are aware of work by Guerini and Savo~\cite{guerini2004eigenvalue},
% who estimate the Poincar\'e--Friedrichs constants over smoothly bounded convex domains in the Hilbert space setting, when $p=2$. 
who provide lower bounds for the spectrum of the Hodge--Laplace operator on convex domains with smooth boundary, % DONE: Theo : smoothly bounded -> with smooth boundary
thus giving upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence for the curl and divergence operators, when $p=2$.
By duality, that implies lower bounds for the case of full boundary conditions.
% They also provide sequences of inequalities, such as the following in 3D smoothly bounded convex domains:
% \begin{align}\label{math:intro:guerinisavo}
%     C_{\grad,\Omega,2} \geq C_{\curl,\Omega,2} \geq C_{\divergence,\Omega,2}.
% \end{align}
% We emphasize that \cite{guerini2004eigenvalue} only addresses smoothly bounded domains, but the extension to convex domains is a technicality.

When the domain is star-shaped with respect to a ball, then estimates for the Poincar\'e--Friedrichs constants 
follow from bounds on the operator norms of regularized Poincar\'e operators~\cite{costabel2010bogovskiui}
as mappings between Lebesgue spaces. 
We are aware of estimates for the higher-order seminorms of these potentials~\cite{guzman2021estimation} (see also~\cite{gallistl2023computational}),
but estimates in Lebesgue norms have not been made explicit in the literature yet, to the best of our knowledge.
\\
% DONE: cite Guzman and Salgado in a reasonable manner

The main objective of this manuscript is bounding the Poincar\'e--Friedrichs constants over local stars within triangulated domains. 
Clearly, not all local stars describe convex subdomains. 
One possible avenue towards Poincar\'e--Friedrichs constants is the observation that local stars in triangulations are star-shaped with respect to a ball,
which enables tools such as averaged Poincar\'e and Bogovski\u{\i} operators~\cite{costabel2010bogovskiui}. 
However, estimates that rely on this geometric condition deteriorate when the aforementioned ball has radius much smaller than the domain diameter. 
This is not as much a problem over patches around interior subsimplices, where the size of the interior ball only depends on the shape regularity of the triangulation. 
But that interior ball can be arbitrarily small when the patch is around a boundary simplex, even if the mesh has good shape regularity: 
this occurs when the domain has sharp reentrant corners, and some illustrative limit cases include the slit domain~\cite{veeser2012poincare} and the crossed bricks domain~\cite{licht2019smoothed}, % DONE: MV : find the examples in the paper by verfurth and veeser 
which contain finite element patches are not star-shaped with respect to any ball.

Domains that admit finite triangulations allow for different techniques to attain Poincar\'e--Friedrichs inequalities. 
However, the majority of results only address such inequalities for scalar-valued functions, not for vector fields. 
We review some of the main outcomes. 
Veeser and Verf\"urth provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for finite element stars~\cite{veeser2012poincare}. Naturally, their estimate deteriorates when the shape regularity of the mesh deteriorates. 
% 
The literature~\cite{Eym_Gal_Her_00,vohralik2005discrete,ern2020stable,ern2021finite,Chaum_Voh_p_rob_3D_H_curl_23,Voh_loc_glob_H1_24} knows different variations of a popular avenue towards Poincar\'e--Friedrichs constants that circumvents the effect of low boundary regularity and constructs potentials step-by-step: 
starting with a single triangle, the potential is constructed over increasing subdomains, until the entire domain is exhausted. That procedure applies to general triangulated domains, not only local stars, though the latter is our main interest here. 
The underlying idea is that we first construct a potential for the gradient over an initial simplex. 
Every time we have found a potential over a subdomain, we construct a potential over a neighboring simplex or patch:
along the interfacing intersection, the two potentials will only differ by a constant, 
and that difference can easily be removed to ensure continuity across that interface. 
% Estimates for Poincar\'e--Friedrichs constants of triangulated domains are known in various forms.

>>>>

\documentclass[10pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\input{header}

@@

\usepackage{bm} % preferred version of bold symbols
\usepackage[normalem]{ulem} % for strikethrough (\sout) in collaborative writing
\usepackage{cancel} % for strikethrough in collaborative writing (math formulas)
\usepackage[textsize=tiny,color=blue!50!white]{todonotes}
\setlength{\marginparwidth}{2cm}

\newcommand\cye[1]{%
  \protect\leavevmode
  \begingroup
    \color{red!35!yellow}%
    #1%
  \endgroup
}

\title{Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over \cye{convex domains and} domains with shellable triangulations}
\author{
    Th\'eophile Chaumont-Frelet\thanks{Affiliation of the First Author, \texttt{email1@example.com}} \and
    Martin Werner Licht\thanks{Affiliation of the Second Author, \texttt{martin.licht@epfl.ch}} \and
    Martin Vohral\'ik\thanks{\cye{Inria, 48 rue Barrault, 75647 Paris, France \&
CERMICS, Ecole des Ponts, 77455 Marne-la-Vallée, France}, \texttt{\cye{martin.vohralik@inria.fr}}}
}

@@

%\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\notice}[1]{\textcolor{red}{REMARK: #1}}

\newcommand{\disk}{\cye{ball}}

@@

\begin{abstract}
    We construct potentials for the gradient, the curl\cye{,} and the divergence operators, and more generally the exterior derivative, over domains with shellable triangulations. This class of triangulations \cye{in particular} includes local \cye{patches (stars)} in dimension two and three. 
    \cye{We derive} computable bounds \cye{on} the operator norms of our potentials that only depend on the geometry. We thus obtain computable upper bounds for \cye{constants in} Poincar\'e--Friedrichs inequalities \cye{as well as} computable lower bounds for the eigenvalues of several vector Laplacians. 
    As a result with independent standing, we \cye{also} establish Poincar\'e--Friedrichs inequalities \cye{with computable constants} for the $L^{p}$ de~Rham complex over convex bounded domains. 
    This is achieved through \cye{a suitable construction of regularized} Poincar\'e and Bogovski\u{\i} potential operators \cye{where we can bound their operator norms}.
    \cye{Computational examples illustrate the theoretical findings.}
\end{abstract}

\tableofcontents

\todo[inline]{MV: Added table of contents: for us, to see the structure of and the main terminology used in the paper}

@@

\todo[inline]{MV: Added subsections and reorganized.}

Potentials for the differential operators of vector calculus and exterior calculus are of fundamental importance. The operator norms of the\cye{se} potentials are upper bounds for the Poincar\'e--Friedrichs constants, which quantify the fundamental stability features of a variety of partial differential equations. In numerical methods, estimates for the Poincar\'e--Friedrichs constant\cye{s} enter the stability and convergence theory and spectral bounds for \cye{stiffness} matrices. \cye{Computable upper bounds on the Poincar\'e--Friedrichs constants also give computable lower bounds for the eigenvalues of the associated Laplacians}. However, while potential operators and Poincar\'e--Friedrichs constants for the gradient have been subject to extensive study, quantifiable results regarding the curl and divergence operators or \cye{more generally} the exterior derivative are \cye{largely unavailable}.

\cye{Our main purpose here is to fill this gap. In the gradient case, we can go over with so-called face-connected triangulations. However, to proceed} in the \cye{general exterior derivative case, we are brought to} so-called shellable triangulations. 
Even though only contractible domains can ever admit a shellable triangulation, 
having computable upper bounds for such domains is an important stepping stone towards more general cases. \cye{Still, we cover most practical nonconvex and non-star-shaped triangulated domains, and as a particular example, local triangulations} around a simplex within a larger triangulation \cye{(the so-called local patches or stars), which are} shellable in two and three spaces dimensions. \cye{Over convex bounded domains, we derive Poincar\'e--Friedrichs inequalities with computable constants for the whole $L^{p}$ de~Rham complex.}

\cye{Let $\Omega \subseteq \bbR^n$, $n \geq 1$, be an connected open set.} \todo[inline]{MV: Here not necessarily bounded? But then we sometimes need $\Omega \subseteq \bbR^{n}$ of finite measure \ldots I would be fine with $\Omega$ bounded all the time.} Given any $1 \leq p \leq \infty$, we let $L^{p}(\Omega)$ denote the Lebesgue space over $\Omega$ with integrability exponent $p$. Then $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ is the corresponding Lebesgue space of vector\cye{-valued} fields.

\subsection{\cye{Potential for the gradient operator and the corresponding Poincar\'e--Friedrichs inequality}} \label{section:intro_grad}

Let $W^{1,p}(\Omega)$ denote the first-order Sobolev space over $\Omega$ with integrability exponent $p$, 
\begin{align} \label{equation:W1p}
    W^{1,p}(\Omega) := \left\{u \in L^{p}(\Omega) \cye{\suchthat} \nabla u \in \bfL^{p}(\Omega) \right\}.
\end{align}
For a vector-valued field $\bff$ that is a distributional gradient of an object from $W^{1,p}(\Omega)$, $\bff \in \nabla W^{1,p}(\Omega)$, we are \cye{interested in} {\em potentials}, i.e., scalar-valued functions $\Phi_{\cye{\grad}}( \bff ) \in W^{1,p}(\Omega)$ such that $\nabla \Phi_{\cye{\grad}}( \bff ) = \bff$. Then, 
%
\begin{align*}
    \Phi_{\cye{\grad}} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)
\end{align*}
is the {\em potential operator}. A particular example of our interest is
%
\begin{subequations}\label{math:pot_grad}\begin{align}\label{math:pot_grad_def}
    \Phi_{\cye{\grad}}( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)}.
\end{align}
%
Actually, in the present gradient setting and since $\Omega$ is connected, already the gradient constraint determines $\Phi_{\cye{\grad}}( \bff )$ up to a constant\cye{, so that~\eqref{math:pot_grad_def} is a one-dimensional minimization problem}.

\cye{We want to see whether the $L^{p}(\Omega)$ norm of the potential $\Phi_{\cye{\grad}}( \bff )$ is bounded in terms of the $\bfL^{p}(\Omega)$ norm of the source field $\bff$, i.e., we want to evaluate the {\em operator norm}
\begin{align}\label{math:pot_grad_norm}
    C_{\grad,\Omega,p} := \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\cye{\grad}}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)} }.
\end{align}\end{subequations}}
This operator norm is indeed finite and it equivalently turns out to be the best constant in the {\em Poincar\'e--Friedrichs inequality}: for every vector field $\bff \in \nabla W^{1,p}(\Omega)$, there exists $u \in W^{1,p}(\Omega)$ such that $\nabla u = \bff$ and 
\begin{subequations}\label{math:PF_gradient}\begin{align} \label{math:PF_gradient_1}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{\cye{\bfL}^{p}(\Omega)}.
\end{align}
Actually, \eqref{math:PF_gradient_1} is further equivalent to the more conventional writing
\begin{align} \label{math:PF_gradient_2}
    \min_{ c \in \bbR } \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p}
    \| \nabla u \|_{\cye{\bfL}^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    .
\end{align}\end{subequations}
We call $C_{\grad,\Omega,p}$ the \emph{Poincar\'e--Friedrichs constant} for the gradient operator with exponent $p$. \cye{A concise writing of all~\eqref{math:pot_grad}--\eqref{math:PF_gradient} is
\begin{align} \label{math:pot_PF_grad}
    \min\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } 
    \| u \|_{L^{p}(\Omega)}
    &\leq 
    C_{\grad,\Omega,p}
    \| \bff \|_{\cye{\bfL}^{p}(\Omega)}.
\end{align}
for all $\bff \in \nabla W^{1,p}(\Omega)$.}

Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities~\eqref{math:pot_PF_grad}. Most research on optimal constants has addressed the case of convex domains. 
Notably, if the constants are required to depend on the convex domain only via its diameter, 
then the optimal constants $C_{\grad,\Omega,p}$ for any $1 \leq p < \infty$ are known explicitly~\cite{Pay_Wei_Poin_conv_60,bebendorf2003note,acosta2004optimal,esposito2013poincare,ferone2012remark}.
When the domain is not convex but star-shaped with respect to a ball,
then polynomial interpolation estimates already imply the Poincar\'e\cye{--Friedrichs} \todo{MV: OK???} inequality~\cite{brenner2008mathematical,ern2021finite}. 
Poincar\'e\cye{--Friedrichs} \todo{MV: OK???} inequalities \cye{actually} hold over \cye{merely}\todo{MV: OK???} star-shaped open sets~\cite[Theorem~3.1]{hurri1988poincare}. \cye{We present a more profound discussion below in Section~\ref{section:poincare}, where we also discuss the difference with the Poincar\'e inequality, where the mean value of $u$ is fixed to zero, and address why we later employ linear potential operators to estimate the Poincar\'e--Friedrichs constant and not directly the minimizing but nonlinear (for $p \neq 2$) potential operators such as~\eqref{math:pot_grad_def}.}

\subsection{\cye{Potentials for the curl and divergence operators and the corresponding Poincar\'e--Friedrichs inequalities}} \label{section:intro_curl_div}

Consider the vector-valued spaces
\begin{subequations}\label{equation:W_curl_div_p} \begin{align}
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) \cye{\suchthat} \curl \bfu \in \bfL^{p}(\Omega) \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in \bfL^{p}(\Omega) \cye{\suchthat} \divergence \bfu \in L^{p}(\Omega) \right\}
    .
\end{align}\end{subequations}
The members of these spaces are those vector fields whose distributional curls and divergences are \cye{respectively} in \cye{the} Lebesgue spaces \cye{$\bfL^{p}$ or $L^{p}(\Omega)$}.  
This only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector\cye{-valued} fields.

\cye{Let a vector-valued field $\bff$ that is a distributional curl of an object from $\bfW^{p}(\curl,\Omega)$, $\bff \in \curl \bfW^{p}(\curl,\Omega)$, and scalar-valued field $f$ that is a distributional divergence of an object from $\bfW^{p}(\divergence,\Omega)$, $f \in \divergence \bfW^{p}(\divergence,\Omega)$, be given.} 
We are interested in constants for inequalities such as 
\begin{subequations}\label{math:pot_PF_curl_div} \begin{align}
    \min\limits_{ \substack{ \cye{\bfu} \in \bfW^{p}(\curl,\Omega) \\ \curl \cye{\bfu} = \bff } } 
    \| \cye{\bfu} \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)} \label{math:pot_PF_curl}
    ,
    \\
    \min\limits_{ \substack{ \cye{\bfu} \in \bfW^{p}(\divergence,\Omega) \\ \divergence \cye{\bfu} = f } } 
    \| \cye{\bfu} \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)} \label{math:pot_PF_div}
    ,
\end{align}\end{subequations}
and how these constants can be estimated by the operator norms of {\em linear} potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
The fundamental difference \cye{with respect to Section~\ref{section:intro_grad}} is that the curl and divergence have infinite-dimensional kernels. In Section~\ref{section:intro_grad}, the kernel of the gradient is the one-dimensional space of constant functions. It is thus trivially complemented for all $p$, with a canonical choice of projection. By contrast, in the Banach space case \cye{$p \neq 2$}, it is not immediately evident that the kernels of the curl and divergence operators are complemented\cye{,} and a canonical projection only exists here in the Hilbert setting $p=2$. \cye{Actually,} the existence of a bounded linear potential operator is \cye{here} a non-trivial fact even in the Hilbert space case \cye{$p = 2$}.

\cye{Let for a moment the topology of $\Omega$ be trivial, let $\Omega$ be weakly Lipschitz, and consider the Hilbert case $p=2$. Then, \eqref{math:pot_PF_curl} is related to the so-called Poincar\'e--Friedrichs--Weber inequality
\begin{align}
    \label{math:Weber}
    \| \bfu\|_{L^{2}(\Omega)} \leq \tilde C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)},
\end{align}
for all $\bfu \in \bfW^{2}(\curl,\Omega) \cap \bfW^{2}(\divergence,\Omega)$ with either the tangential or the normal trace zero on $\partial \Omega$ and $\divergence \bfu = 0$. 
Yet another writing is 
\label{math:PF_orth} \begin{align}
    \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)}, \label{math:PF_orth_curl}\\
    \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\divergence,\Omega,p} \| \divergence \bfu \|_{L^{2}(\Omega)} \label{math:PF_orth_div}
\end{align}
for respectively all functions $\bfu$ from $\bfW^{2}(\curl,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv$ from $\bfW^{2}(\curl,\Omega)$ with $\curl \bfv = 0$ and all functions $\bfu$ from $\bfW^{2}(\divergence,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv$ from $\bfW^{2}(\divergence,\Omega)$ with $\divergence \bfv = 0$. We refer to Friedrichs~\cite[equation~(5)]{Fried_diff_forms_55}, Gaffney~\cite[equation~(2)]{Gaff_Hilbert_harm_55}, Weber~\cite{Web_compact_Maxw_80}, \cite[Lemmas~3.4 and~3.6]{Gir_Rav_NS_86}, \cite[Proposition~7.4]{Fer_Gil_Maxw_BC_97}, and the references therein. Still} in the Hilbert space case $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space with \cye{homogeneous} Dirichlet boundary conditions \cye{on $\partial \Omega$} implies, via duality~\eqref{math:pot_PF_div}. However, that easy observation seems restricted to $p=2$.

Not much attention\cye{, however,} seems to have been given to the study of \cye{computable} constants in~\eqref{math:pot_PF_curl_div}, even if the domain $\Omega$ is convex. \cye{Important results in the Hilbert case $p=2$ can be found in~\cite{guerini2004eigenvalue, Paul_Vald_PF_grad_curl_div_20} that we discuss now in the broder setting of} the exterior calculus formalism\cye{, but the Banach case $p \neq 2$ seems largely unexplored.}

\subsection{\cye{Potentials for the exterior derivative and the corresponding Poincar\'e--Fried\-richs inequalities}} \label{section:intro_EC}

Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the \cye{space} dimension and highlights \cye{extends the} underlying geometric structures common to the gradient, curl, and divergence operators \cye{in three space dimensions}.
\cye{In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}: for every $w \in W^{p}\Alt^{k}(\Omega)$, there exists $u \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ and such that (all the notation is in detail fixed in Section~\ref{section:calculus} below)
\begin{align} \label{math:PF_DF} 
    \| u \|_{L^{p}(\Omega} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
}\todo[inline]{MV: The roles of $u$ and $w$ are flipped in the various parts of the paper, see~\eqref{math:PF_DF} in comparsion with Lemma~\ref{lemma:PFexteriorsimplex}. It would be great to unify.}

In the Hilbert space case $p=2$, Guerini and Savo~\cite{guerini2004eigenvalue} provide lower bounds for the spectrum of the Hodge--Laplace operator on convex domains with smooth boundary, thus giving upper bounds for the Poincar\'e--Friedrichs constant of the exterior derivative, and hence \cye{in particular} for the curl and divergence operators.
\cye{A comparison result} shows that the Poincar\'e--Friedrichs constant for the gradient already estimates the corresponding constants for the curl and divergence operators. By duality, this also \cye{gives} upper estimate\cye{s} of the Poincar\'e--Friedrichs constants for the gradient, curl, and divergence \cye{operators} subject to Dirichlet, tangential, and normal boundary conditions, respectively, along the entire boundary. 
\cye{Computable constants for Lipschitz domains have recently been addressed in Pauly and Valdman~\cite{Paul_Vald_PF_grad_curl_div_20}, see also the references therein.}
